---
title: "Correlation Between QT/QTc Interval and Calcium"
author: "JW Yi"
highlighter: highlight.js
output: word_document
job: null
knit: slidify::knit2slides
mode: selfcontained
hitheme: tomorrow
subtitle: null
framework: io2012
widgets: bootsrtap
---


## Calcium, iCalcium, PTH, P vs QT/QTc 

- For developing calcium level prediction model from the ECG wave, we initially analyzed correlation between lab data and ECG data, used in SNUH.

---

## Data Preperation

```{r}  
qtdata <- read.delim("C://Work/QTcalcium/QT data.txt", header=T, na.strings="NA")
dim(qtdata)
head(qtdata)
```  

---   

## Pairs Plot  

```{r, echo=FALSE}  
pairs(qtdata, col="blue")
```  

---   

## QT Intreval according to Calcium   

```{r}
Ca_QT_Fit <- lm(QT ~ Calcium, data=qtdata)
summary(Ca_QT_Fit)
```

---     

## QT Intreval according to Ionized Calcium     

```{r}
iCa_QT_Fit <- lm(QT ~ iCalcium, data=qtdata)
summary(iCa_QT_Fit)
```

---   

## QT Intreval according to PTH    

```{r}
PTH_QT_Fit <- lm(QT ~ PTH, data=qtdata)
summary(PTH_QT_Fit)
```

---   

## QT Intreval according to Phosphate  

```{r}
P_QT_Fit <- lm(QT ~ Phosphate, data=qtdata)
summary(P_QT_Fit)
```

---   
  
## QTc Intreval according to Calcium   

```{r}
Ca_QTc_Fit <- lm(QTc ~ Calcium, data=qtdata)
summary(Ca_QTc_Fit)
```

---      

## QTc Intreval according to Ionized Calcium   

```{r}
iCa_QTc_Fit <- lm(QTc ~ iCalcium, data=qtdata)
summary(iCa_QTc_Fit)
```
---   

## QTc Intreval according to PTH    

```{r}
PTH_QTc_Fit <- lm(QTc ~ PTH, data=qtdata)
summary(PTH_QTc_Fit)
```

---   

## QTc Intreval according to Phosphate    

```{r}
P_QTc_Fit <- lm(QTc ~ Phosphate, data=qtdata)
summary(P_QTc_Fit)
```

---     

## Correlation Plots    

```{r, echo=FALSE}
library(ggplot2)
library(gridExtra)
g1 <- ggplot(qtdata, aes(Calcium, QT))
g1 <- g1 + geom_point(color="blue", size=3, alpha=1/2) +
    geom_smooth(method="lm") +
    labs(title="QT interval and Calcium", 
         x="Calcium", y="QT interval")

g2 <- ggplot(qtdata, aes(iCalcium, QT))
g2 <- g2 + geom_point(color="blue", size=3, alpha=1/2) +
    geom_smooth(method="lm") +
    labs(title="QT interval and ionized Calcium", 
         x="Ionized Calcium", y="QT interval")

g3 <- ggplot(qtdata, aes(Calcium, QTc))
g3 <- g3 + geom_point(color="red", size=3, alpha=1/2) +
    geom_smooth(method="lm") +
    labs(title="QTc interval and Calcium", 
         x="Calcium", y="QTc interval")

g4 <- ggplot(qtdata, aes(iCalcium, QTc))
g4 <- g4 + geom_point(color="red", size=3, alpha=1/2) +
    geom_smooth(method="lm") +
    labs(title="QTc interval and ionized Calcium", 
         x="Ionized Calcium", y="QTc interval")

grid.arrange(g1, g2, g3, g4)
```

---   

## Further Tasks      
 
- Collect more data for making prediction model. 
- Data seperation into training and testing samples.
- Make prediction model for estimate calcium level from the ECG data.
- Apply this model on real ECG data.

---    

