---
title: "Estimating Calcium Level From QT Interval"
author: "JW Yi"
date: "Sunday, December 14, 2014"
output: beamer_presentation
---

## Estimating Calcium Level From QT/QTc Interval

- Hypocalcemia is an important post operative complication after thyroid surgery.
- Still, we can only measure calcium level by patient's blood.
- Acute hypocalcemia causes prolongation of the QT interval.
- Can we infer the blood calcium level from the QT interval?

## Load prepaired data.
```{r}
patients <- read.csv("D://Work/QTcalcium/20141214/QT calcium.csv")
head(patients, 3)
```

## Convert initial data into long format.
```{r}
qtdata <- read.csv("D://Work/QTcalcium/20141214/QT calcium Long Format.csv")
head(qtdata)
```

## Summary Statistics.
```{r}
summary(qtdata)
library(caret); library(gridExtra)
```

## Analysis From Machine Data

- ECG machine automatically displays QT and QTc interval. 
- Can we trust this interval?

## QT/QTc From ECG Machine VS Electrolytes.
```{r, echo=FALSE, fig.width=8, fig.height=8}
featurePlot(x = qtdata[,2:3], y = qtdata[,7:10], plot="pairs")
```

## Linear Regression Analysis, Machine QT vs Calcium
```{r}
Ca_QT_M_Fit <- lm(QT_M ~ Ca, data=qtdata)
summary(Ca_QT_M_Fit)
```

## Linear Regression Analysis, Machine QT vs Ionized Calcium
```{r}
iCa_QT_M_Fit <- lm(QT_M ~ iCa, data=qtdata)
summary(iCa_QT_M_Fit)
```

## Linear Regression Analysis, Machine QTc vs Calcium
```{r}
Ca_QTc_M_Fit <- lm(QTc_M ~ Ca, data=qtdata)
summary(Ca_QTc_M_Fit)
```

## Linear Regression Analysis, Machine QTc vs Ionized Calcium
```{r}
iCa_QTc_M_Fit <- lm(QTc_M ~ iCa, data=qtdata)
summary(iCa_QTc_M_Fit)
```

## Relationship between Machine QT/QTc and Ca/iCa 

- Have Significant p-values (under 0.05) but slopes were positive.
- This is the opposing result from the basic calcium vs QT interval therory. 
- We cannot trust QT/QTc from the ECG machine.

## Analysis From Cardiologist's data

- Cardiologist have re-calculated QT and QTc interval manually. 
- Can we trust him? 

## QT/QTc From Cardiologists VS Electrolytes.
```{r, echo=FALSE, fig.width=8, fig.height=8}
featurePlot(x = qtdata[,4:5], y = qtdata[,7:10], plot="pairs")
```

## Linear Regression Analysis, Cardiologist's QT vs Calcium
```{r}
Ca_QT_C_Fit <- lm(QT_C ~ Ca, data=qtdata)
summary(Ca_QT_C_Fit)
```

## Linear Regression Analysis, Cardiologist's QT vs Ionized Calcium
```{r}
iCa_QT_C_Fit <- lm(QT_C ~ iCa, data=qtdata)
summary(iCa_QT_C_Fit)
```

## Linear Regression Analysis, Cardiologist's QTc vs Calcium
```{r}
Ca_QTc_C_Fit <- lm(QTc_C ~ Ca, data=qtdata)
summary(Ca_QTc_C_Fit)
```

## Linear Regression Analysis, Cardiologist's QTc vs Ionized Calcium
```{r}
iCa_QTc_C_Fit <- lm(QTc_C ~ iCa, data=qtdata)
summary(iCa_QTc_C_Fit)
```

## Correlation Plot
```{r, echo=FALSE, warning=FALSE, fig.width=8, fig.height=8}
g1 <- ggplot(qtdata, aes(QT_C, Ca))
g1 <- g1 + geom_point(color="blue", size=3, alpha=1/2) +
    geom_smooth(method="lm") +
    labs(title="QT vs Calcium\np-value=0.0191\nslope=-2.391", 
         x="QT interval", y="Calcium")

g2 <- ggplot(qtdata, aes(QT_C, iCa))
g2 <- g2 + geom_point(color="blue", size=3, alpha=1/2) +
    geom_smooth(method="lm") +
    labs(title="QT vs ionized Calcium\np-value=0.0276\nslope=-2.244", 
         x="QT interval", y="Ionized Calcium")

g3 <- ggplot(qtdata, aes(QTc_C, Ca))
g3 <- g3 + geom_point(color="red", size=3, alpha=1/2) +
    geom_smooth(method="lm") +
    labs(title="QTc vs Calcium\np-value=2.08e-05\nslope=-4.521", 
         x="QTc interval", y="Calcium")

g4 <- ggplot(qtdata, aes(QTc_C, iCa))
g4 <- g4 + geom_point(color="red", size=3, alpha=1/2) +
    geom_smooth(method="lm") +
    labs(title="QTc vs Ionized Calcium\np-value=3.6e-07\nslope=-5.544", 
         x="QTc interval", y="Ionized Calcium")

grid.arrange(g1, g2, g3, g4)
```

## Relationship between Cardiologist's QT/QTc and Ca/iCa 

- Have Significant p-values (under 0.05) and negative slopes.
- This result is appropriate to basic calcium vs QT interval therory. 
- Cardiologit's interperations were amazing !


## Changing QT/QTc Interval and Ca/iCa in each patients.

- We need to check the changing QT/QTc and Calcium/ionized Calcium level during preop, POD#1 and POD#2.
- We applied paired t-test and ANOVA among the three days in each patients.

## QT interval change in each patients, Preop vs POD#1
```{r}
t.test(patients$QT_C_Pre, patients$QT_C_POD1, paired=TRUE)
```

## QT interval change in each patients, Preop vs POD#2
```{r}
t.test(patients$QT_C_Pre, patients$QT_C_POD2, paired=TRUE)
```

## QT interval change in each patients, POD#1 vs POD#2
```{r}
t.test(patients$QT_C_POD1, patients$QT_C_POD2, paired=TRUE)
```

## QT interval changes in each patients, ANOVA 
```{r}
qtdata2 <- read.csv("D://Work/QTcalcium/20141214/QT calcium for ANOVA.csv")
aov.out1 <- aov(QT_C ~ Day, data=qtdata2)
TukeyHSD(aov.out1)
```

## QTc interval change in each patients, Preop vs POD#1
```{r}
t.test(patients$QTc_C_Pre, patients$QTc_C_POD1, paired=TRUE)
```

## QTc interval change in each patients, Preop vs POD#2
```{r}
t.test(patients$QTc_C_Pre, patients$QTc_C_POD2, paired=TRUE)
```

## QTc interval change in each patients, POD#1 vs POD#2
```{r}
t.test(patients$QTc_C_POD1, patients$QTc_C_POD2, paired=TRUE)
```

## QTc interval changes in each patients, ANOVA test
```{r}
aov.out2 <- aov(QTc_C ~ Day, data=qtdata2)
TukeyHSD(aov.out2)
```

## Post-Hoc Analysis, QTc interval changes
```{r}
plot(TukeyHSD(aov.out2))
```

## Caicium level change in each patients, Preop vs POD#1
```{r}
t.test(patients$Ca_Pre, patients$Ca_POD1, paired=TRUE)
```

## Caicium level change in each patients, Preop vs POD#2
```{r}
t.test(patients$Ca_Pre, patients$Ca_POD2, paired=TRUE)
```

## Caicium level change in each patients, POD#1 vs POD#2
```{r}
t.test(patients$Ca_POD1, patients$Ca_POD2, paired=TRUE)
```

## Calcium level changes in each patients, ANOVA test
```{r}
aov.out3 <- aov(Ca ~ Day, data=qtdata2)
TukeyHSD(aov.out3)
```

## Post-Hoc Analysis, Calcium level changes
```{r}
plot(TukeyHSD(aov.out3))
```

## Ionized Caicium level change in each patients, Preop vs POD#1
```{r}
t.test(patients$iCa_Pre, patients$iCa_POD1, paired=TRUE)
```

## Ionized Caicium level change in each patients, Preop vs POD#2
```{r}
t.test(patients$iCa_Pre, patients$iCa_POD2, paired=TRUE)
```

## Ionized Caicium level change in each patients, POD#1 vs POD#2
```{r}
t.test(patients$iCa_POD1, patients$iCa_POD2, paired=TRUE)
```

## Ionized Calcium level changes in each patients, ANOVA test
```{r}
aov.out4 <- aov(iCa ~ Day, data=qtdata2)
TukeyHSD(aov.out4)
```

## Post-Hoc Analysis, Ionized Calcium level changes
```{r}
plot(TukeyHSD(aov.out4))
```

## Summary 

- There were significant QTc interval change in Preop vs POD#1 / Preop vs POD#2.
- There were significant Ca and iCa change in Preop vs POD#1 / Preop vs POD#2.
- Using QTc interval is appropriate for estimating Ca/iCa level. 
