---
title: "FNA result according to tumor size"
author: "JW Yi"
output: html_document
---
# FNA data preperation
```{r,echo=FALSE}
library(epiR)
library(epitools)
library(gmodels)
fna <- read.table("D://Work/FNA.txt", header=T, sep='\t')
fna$Category <- factor(fna$FNA, levels=c(2,3,4,5,6), labels = c("Benign","AUS","FN","Suspicious","Malignant"))
summary(fna)
attach(fna)
```

# Size Distribution
```{r}
Size.cat <- function(x, lower=0, upper, by=1, sep="-", above.char=">") {
     labs <- c(paste(seq(lower, upper - by , by=by), 
                     seq(lower + by , upper , by=by),
                     sep=sep),
               paste(upper, above.char, sep=""))
     cut(x, breaks=c(seq(lower, upper, by=by), Inf), right=FALSE, labels=labs)
}

Size.freq <- table(Size.cat(fna$Size, upper=4))

library(RColorBrewer)
col <- brewer.pal(5, "Set1")

b <- barplot(Size.freq, main="Size Distribution", col=col, xlab = "Tumor Size", ylab="Number of Observations", ylim=c(0,100))
text(x = b, y = Size.freq, labels= c(83, 90, 47, 16, 15), pos=3, cex=1.0, col="red")
box()
```

# Cross Table Analysis
## Data Preperation
```{r}
FNA.binary4 <- ifelse(fna$FNA >= 4, "Malignant", "Benign")
FNA.binary4 <- as.factor(FNA.binary4)
FNA.binary5 <- ifelse(fna$FNA >= 5, "Malignant", "Benign")
FNA.binary5 <- as.factor(FNA.binary5)
FNA.binary6 <- ifelse(fna$FNA >= 6, "Malignant", "Benign")
FNA.binary6 <- as.factor(FNA.binary6)

Size.binary1 <- ifelse(fna$Size <= 1, "< 1cm", ">= 1cm")
Size.binary1 <- as.factor(Size.binary1)
Size.binary2 <- ifelse(fna$Size <= 2, "< 2cm", ">= 2cm")
Size.binary2 <- as.factor(Size.binary2)
Size.binary3 <- ifelse(fna$Size <= 3, "< 3cm", ">= 3cm")
Size.binary3 <- as.factor(Size.binary3)

summary(FNA.binary4)
summary(FNA.binary5)
summary(FNA.binary6)
summary(Size.binary1)
summary(Size.binary2)
summary(Size.binary3)
```

## FNA borderline : 4, Size 1cm, 2cm, 3cm
```{r}
t.4.1 <- table (FNA.binary4, Size.binary1)
t.4.1
oddsratio(t.4.1)
```
OR 0.774, p-value 0.379
```{r}
t.4.2 <- table(FNA.binary4, Size.binary2)
t.4.2
oddsratio(t.4.2)
```
OR 0.764, p-value 0.383
```{r}
t.4.3 <- table(FNA.binary4, Size.binary3)
t.4.3
oddsratio(t.4.3)
```
OR 0.293, p-value 0.002 **

## FNA borderline : 5, Size 1cm, 2cm, 3cm
```{r}
t.5.1 <- table(FNA.binary5, Size.binary1)
t.5.1
oddsratio(t.5.1)
```
OR 0.570, p-value 0.032 **
```{r}
t.5.2 <- table(FNA.binary5, Size.binary2)
t.5.2
oddsratio(t.5.2)
```
OR 0.485, p-value 0.012 **
```{r}
t.5.3 <- table(FNA.binary5, Size.binary3)
t.5.3
oddsratio(t.5.3)
```
OR 0.273, p-value 0.004 ** 

## FNA borderline : 6, Size 1cm, 2cm, 3cm
```{r}
t.6.1 <- table(FNA.binary6, Size.binary1)
t.6.1
oddsratio(t.6.1)
```
OR 0.432, p-value 0.009 **
```{r}
t.6.2 <- table(FNA.binary6, Size.binary2)
t.6.2
oddsratio(t.6.2)
```
OR 0.688, p-value 0.320 
```{r}
t.6.3 <- table(FNA.binary6, Size.binary3)
t.6.3
oddsratio(t.6.3)
```
OR 0.792, p-value 0.794 (Fisher)

## Summary
```{r}
a <- c(oddsratio(t.4.1)[[3]][6], oddsratio(t.4.1)[[2]][2], oddsratio(t.4.1)[[2]][4],oddsratio(t.4.1)[[2]][6])
b <- c(oddsratio(t.4.2)[[3]][6], oddsratio(t.4.2)[[2]][2], oddsratio(t.4.2)[[2]][4],oddsratio(t.4.1)[[2]][6])
c <- c(oddsratio(t.4.3)[[3]][6], oddsratio(t.4.3)[[2]][2], oddsratio(t.4.3)[[2]][4],oddsratio(t.4.1)[[2]][6])
d <- c(oddsratio(t.5.1)[[3]][6], oddsratio(t.5.1)[[2]][2], oddsratio(t.5.1)[[2]][4],oddsratio(t.5.1)[[2]][6])
e <- c(oddsratio(t.5.2)[[3]][6], oddsratio(t.5.2)[[2]][2], oddsratio(t.5.2)[[2]][4],oddsratio(t.5.1)[[2]][6])
f <- c(oddsratio(t.5.3)[[3]][6], oddsratio(t.5.3)[[2]][2], oddsratio(t.5.3)[[2]][4],oddsratio(t.5.1)[[2]][6])
g <- c(oddsratio(t.6.1)[[3]][6], oddsratio(t.6.1)[[2]][2], oddsratio(t.6.1)[[2]][4],oddsratio(t.6.1)[[2]][6])
h <- c(oddsratio(t.6.2)[[3]][6], oddsratio(t.6.2)[[2]][2], oddsratio(t.6.2)[[2]][4],oddsratio(t.6.1)[[2]][6])
i <- c(oddsratio(t.6.3)[[3]][4], oddsratio(t.6.3)[[2]][2], oddsratio(t.6.3)[[2]][4],oddsratio(t.6.1)[[2]][6]) # Fisher
result.final1 <- rbind (a,b,c,d,e,f,g,h,i)
colnames(result.final1) <- c("p-value", "Odds Ratio", "CI upper", "CI lower")
rownames(result.final1) <- c("FNA 4, Size 1", "FNA 4, Size 2", "FNA 4, Size 3*", 
                            "FNA 5, Size 1*", "FNA 5, Size 2*", "FNA 5, Size 3*",
                            "FNA 6, Size 1*", "FNA 6, Size 2", "FNA 6, Size 3")
result.final1
```

# Simple Linear Regression
```{r}
fit1 <- lm(FNA ~ Size, data=fna)
summary(fit1)

plot(Size, FNA, main = "Correlation Between \n Tumor Size and FNA Caterogy", pch=20, col="blue")
abline(fit1, col="red")
text(5, 2.5, "Coefficients = -0.229\n p-value=0.000608")
```

# Logistic Regression
## Binary FNA vs Continuous Size
```{r}
fit4 <- glm(FNA.binary4 ~ fna$Size, family="binomial")
summary(fit4)
exp(fit4$coeff) # odds ratio
exp(confint(fit4))  # Conf. interval

pval <- summary(fit4)$coefficients[8]
or <- exp(fit4$coeff)[[2]]
cilow <- exp(confint(fit4))[[2]]
ciup <- exp(confint(fit4))[[4]]
result.4 <- cbind(pval, or, cilow, ciup)

fit5 <- glm(FNA.binary5 ~ fna$Size, family="binomial")
summary(fit5)
exp(fit5$coeff) # odds ratio
exp(confint(fit5))  # Conf. interval

pval <- summary(fit5)$coefficients[8]
or <- exp(fit5$coeff)[[2]]
cilow <- exp(confint(fit5))[[2]]
ciup <- exp(confint(fit5))[[4]]
result.5 <- cbind(pval, or, cilow, ciup)

fit6 <- glm(FNA.binary6 ~ fna$Size, family="binomial")
summary(fit6)
exp(fit6$coeff) # odds ratio
exp(confint(fit6))  # Conf. interval

pval <- summary(fit6)$coefficients[8]
or <- exp(fit6$coeff)[[2]]
cilow <- exp(confint(fit6))[[2]]
ciup <- exp(confint(fit6))[[4]]
result.6 <- cbind(pval, or, cilow, ciup)

result.final3 <- rbind(result.4, result.5, result.6)
rownames(result.final3) <- c("FNA4*","FNA5*","FNA6")
colnames(result.final3) <- c("p-value","Odds Ratio","CI lower", "CI upper")

```

## Binary FNA vs Binary Size
```{r}
fit4.1 <- glm(FNA.binary4 ~ Size.binary1, family="binomial")
summary(fit4.1)
exp(fit4.1$coeff) # odds ratio
exp(confint(fit4.1))  # Conf. interval

pval <- summary(fit4.1)$coefficients[8]
or <- exp(fit4.1$coeff)[[2]]
cilow <- exp(confint(fit4.1))[[2]]
ciup <- exp(confint(fit4.1))[[4]]
result.4.1 <- cbind(pval, or, cilow, ciup)

fit4.2 <- glm(FNA.binary4 ~ Size.binary2, family="binomial")
summary(fit4.2)
exp(fit4.2$coeff) # odds ratio
exp(confint(fit4.2))  # Conf. interval

pval <- summary(fit4.2)$coefficients[8]
or <- exp(fit4.2$coeff)[[2]]
cilow <- exp(confint(fit4.2))[[2]]
ciup <- exp(confint(fit4.2))[[4]]
result.4.2 <- cbind(pval, or, cilow, ciup)

fit4.3 <- glm(FNA.binary4 ~ Size.binary3, family="binomial")
summary(fit4.3)
exp(fit4.3$coeff) # odds ratio
exp(confint(fit4.3))  # Conf. interval

pval <- summary(fit4.3)$coefficients[8]
or <- exp(fit4.3$coeff)[[2]]
cilow <- exp(confint(fit4.3))[[2]]
ciup <- exp(confint(fit4.3))[[4]]
result.4.3 <- cbind(pval, or, cilow, ciup)

fit5.1 <- glm(FNA.binary5 ~ Size.binary1, family="binomial")
summary(fit5.1)
exp(fit5.1$coeff) # odds ratio
exp(confint(fit5.1))  # Conf. interval

pval <- summary(fit5.1)$coefficients[8]
or <- exp(fit5.1$coeff)[[2]]
cilow <- exp(confint(fit5.1))[[2]]
ciup <- exp(confint(fit5.1))[[4]]
result.5.1 <- cbind(pval, or, cilow, ciup)

fit5.2 <- glm(FNA.binary5 ~ Size.binary2, family="binomial")
summary(fit5.2)
exp(fit5.2$coeff) # odds ratio
exp(confint(fit5.2))  # Conf. interval

pval <- summary(fit5.2)$coefficients[8]
or <- exp(fit5.2$coeff)[[2]]
cilow <- exp(confint(fit5.2))[[2]]
ciup <- exp(confint(fit5.2))[[4]]
result.5.2 <- cbind(pval, or, cilow, ciup)

fit5.3 <- glm(FNA.binary5 ~ Size.binary3, family="binomial")
summary(fit5.3)
exp(fit5.3$coeff) # odds ratio
exp(confint(fit5.3))  # Conf. interval

pval <- summary(fit5.3)$coefficients[8]
or <- exp(fit5.3$coeff)[[2]]
cilow <- exp(confint(fit5.3))[[2]]
ciup <- exp(confint(fit5.3))[[4]]
result.5.3 <- cbind(pval, or, cilow, ciup)

fit6.1 <- glm(FNA.binary6 ~ Size.binary1, family="binomial")
summary(fit6.1)
exp(fit6.1$coeff) # odds ratio
exp(confint(fit6.1))  # Conf. interval

pval <- summary(fit6.1)$coefficients[8]
or <- exp(fit6.1$coeff)[[2]]
cilow <- exp(confint(fit6.1))[[2]]
ciup <- exp(confint(fit6.1))[[4]]
result.6.1 <- cbind(pval, or, cilow, ciup)

fit6.2 <- glm(FNA.binary6 ~ Size.binary2, family="binomial")
summary(fit6.2)
exp(fit6.2$coeff) # odds ratio
exp(confint(fit6.2))  # Conf. interval

pval <- summary(fit6.2)$coefficients[8]
or <- exp(fit6.2$coeff)[[2]]
cilow <- exp(confint(fit6.2))[[2]]
ciup <- exp(confint(fit6.2))[[4]]
result.6.2 <- cbind(pval, or, cilow, ciup)

fit6.3 <- glm(FNA.binary6 ~ Size.binary3, family="binomial")
summary(fit6.3)
exp(fit6.3$coeff) # odds ratio
exp(confint(fit6.3))  # Conf. interval

pval <- summary(fit6.3)$coefficients[8]
or <- exp(fit6.3$coeff)[[2]]
cilow <- exp(confint(fit6.3))[[2]]
ciup <- exp(confint(fit6.3))[[4]]
result.6.3 <- cbind(pval, or, cilow, ciup)

result.final2 <- rbind(result.4.1, result.4.2, result.4.3, result.5.1, result.5.2, result.5.3, 
                      result.6.1, result.6.2, result.6.3)
rownames(result.final2) <- c("FNA 4, Size 1", "FNA 4, Size 2", "FNA 4, Size 3*", 
                            "FNA 5, Size 1*", "FNA 5, Size 2*", "FNA 5, Size 3*",
                            "FNA 6, Size 1*", "FNA 6, Size 2", "FNA 6, Size 3")
colnames(result.final2) <- c("p-value", "Odds Ratio", "CI lower", "CI upper")
result.final2
```

# Final Result
## Chi-square analysis
```{r}
result.final1
```

## Logistric Regression (Binary FNA, Binary Size)
```{r}
result.final2
```

## Linear Regression (All FNA, Continuous Size)
```{r}
plot(Size, FNA, main = "Correlation Between \n Tumor Size and FNA Caterogy", pch=20, col="blue")
abline(fit1, col="red")
text(5, 2.5, "Coefficients = -0.229\n p-value=0.000608")
```

## Logistic Regression (Binary FNA, Continuous Size)
```{r}
result.final3
```

