---
title: "Diagnostic Accuracy of FNA for FVPTC"
author: "JW Yi"
date: "Tuesday, November 04, 2014"
output: html_document
---

# Read dataset and load libraries.

```{r, echo=FALSE, results='hide'}
library(reshape2); library(dplyr); library(epiR); library(epitools); library(gmodels); library(RColorBrewer)
dat <- readRDS("D://Work/FNA/FNA_data.rds")
fna.df <- tbl_df(dat)
fna.df
```

# Summary Statistics according to FNA binary results.

```{r}
Benign_group <- filter(fna.df, FNA_binary == "Benign(2,3,4)")
Malignant_group <- filter(fna.df, FNA_binary == "Malignant(5,6)")
```

* FNA 2,3,4 group summary  
```{r}
summary(Benign_group)
```

* FNA 5,6 group summary   
```{r}
summary(Malignant_group)
```
    
# Linear Regression Model
## Univariable Linear Regression
```{r}
slr1 <- lm(FNA ~ Sex, data=fna.df)
round(summary(slr1)$coef, 3)
slr2 <- lm(FNA ~ Age, data=fna.df)
round(summary(slr2)$coef, 3)
slr3 <- lm(FNA ~ USG, data=fna.df)
round(summary(slr3)$coef, 3)
slr4 <- lm(FNA ~ Size, data=fna.df)
round(summary(slr4)$coef, 3)
```

## Multivariable Linear Regression
```{r}
fit1 <- lm(FNA ~ Sex + Age + USG + Size, data=fna.df)
round(summary(fit1)$coef, 3)
step(fit1)
fit1.reduced <- lm(FNA ~ USG + Size, data=fna.df)
round(summary(fit1.reduced)$coef, 3)
```

# Logistic Regression with binary FNA category
## Univariable Logistic Regression
```{r}
logit1 <- glm(FNA_binary ~ Sex, family="binomial", data=fna.df)
round(summary(logit1)$coef, 3)
logit2 <- glm(FNA_binary ~ Age, family="binomial", data=fna.df)
round(summary(logit2)$coef, 3)
logit3 <- glm(FNA_binary ~ USG, family="binomial", data=fna.df)
round(summary(logit3)$coef, 3)
logit4 <- glm(FNA_binary ~ Size, family="binomial", data=fna.df)
round(summary(logit4)$coef, 3)
```

## Multivariable Logistic Regression
```{r}
logit.all <- glm(FNA_binary ~ Sex + Age + USG + Size, family="binomial", data=fna.df)
summary(logit.all)
step(logit.all)
logit.reduced <- glm(FNA_binary ~ Sex + USG + Size, family = "binomial", data = fna.df)
summary(logit.reduced)
cbind(OR = exp(logit.reduced$coef), exp(confint(logit.reduced)))
```

# Plot

* Barplot  

```{r, echo=FALSE}
col <- brewer.pal(5, "Set1")
tx <- table(fna.df$FNA_binary, fna.df$Size_range)
b <- barplot(tx, legend.text = rownames(tx), col=col[2:1], ylim=c(0, 120))
text(0.7, 25, "51")
text(0.7, 80, "64")
text(1.9, 25, "52")
text(1.9, 75, "45")
text(3.1, 10, "20")
text(3.1, 30, "18")
text(4.3, 10, "18")
text(4.3, 30, "5")
box()
```

* Dotchart   

```{r, echo=FALSE}
dotchart(tx, col=col[2:1])
```

* Stripchart   

```{r, echo=FALSE}
stripchart(Size ~ FNA, method="jitter", vertical=TRUE, data=fna.df, pch=16, col=col, xlab="FNA")
abline(slr4, col="salmon")
```
